Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre

### 2024-10-25 manuscript/manuscript.Rmd
Adding statistical methods used

All statistics were run in R version 4.4.0. We used multivariate statistics provided by the vegan package (version 2.6-6.1, @R-vegan) to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses. Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible. Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995]. For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001]. Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable. Where time of day was included as an explanatory factor in the PERMANOVA, it was binned into 6-hour increments starting from 3AM (e.g. Midday contains casts collected between 9AM and 3PM). Marginal effects are reported for the PERMANOVA analyses to avoid influencing the results with the order of the terms in the model formula.

@@

The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6). Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (R$^2_{depth} = 0.251$, p-value << 0.001). We saw no difference between the 6AM and 6PM samples (R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).

@@

When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy (R$^2$=0.26) and a lower p-value (0.013), while the DCM samples were no longer likely to be distinct between the two eddies (p-value = 0.171) (Figure 6, Supplemental Figure 5).

>>>>

All statistics were run in R version 4.4.0. We used multivariate statistics provided by the vegan package (version 2.6-6.1, @R-vegan) to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses. Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible. Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995]. For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001]. Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable. Where time of day was included as an explanatory factor in the PERMANOVA, it was binned into 6-hour increments starting from 3AM (e.g. Midday contains casts collected between 9AM and 3PM). Marginal effects are reported for the PERMANOVA analyses to avoid influencing the results with the order of the terms in the model formula. Total sample size (*n*) for the eddy transect was 33 per depth (3 biological triplicates at each of 11 stations), while the eddy centers had an *n* of 15 per eddy (3 biological replicates at 5 depths) and the HEE data had an *n* of 24 (3 biological replicates at 4 stations (2 in the cyclone and 2 in the anticyclone) at 2 depths each).

@@

The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6). Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001). We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 4).

@@

When analyzed separately as distinct depths, the 25 meter samples had higher variances explained by eddy and a lower p-value (PERMANOVA, R$^2$=0.26, p-value=0.013), while the DCM samples were no longer likely to be distinct between the two eddies (PERMANOVA p-value = 0.171) (Figure 6, Supplemental Figure 5).
